---
title: "problemTwo"
author: "Daniel Emaasit"
date: "February 24, 2015"
output: html_document
---

# Problem C17.5
Using the data in FERTIL1.RAW. 

## Part (i)
Estimate a poisson regression model for kids, using the same variables in TABLE 13.1. Interpret the coefficient on y82.

```{r}
# read the data into R
library(haven)
fertile_data<-read_dta("FERTIL1.dta")


# estimate a poisson regression model
library(MASS)
model1<-glm(kids~., family="poisson", data=fertile_data)
library(stargazer)
stargazer(model1)

````

## Part (ii)
What is the estimated percentage difference in fertility between a black woman and 
a non black woman, holding other factors fixed?

```{r}


````

## Part (iii)
Obtain the estimated standard deviation Is there evidence of over or underdispersion?

```{r}


````

## Part (iv)
Compute the fitted values from the Posson regression and obtain the R-squared as the squared
correlation betweein kids and estimated kids. Compare this with the R-squared for the linear regression model

```{r}


````

# Problem C17.9
Use the data in APPLE.RAW for this exercise. 

## Part (i)
Of the 660 families in the sample, how many report wanting none of the eco-labeled
apples at the set price?

```{r}
# read the data into R
apple_data<-read_dta("APPLE.dta")

# count the no of families wanting none of the ecolabeled apples
library(dplyr)
apple_data %>% filter(ecolbs==0) %>% count()
````

248 families report wanting none of the eco-labelled apples at the set price

## Part (ii)
Does the variable ecolbs seem to have a continous distribution over strictly positive
values? What implications does your answer have for the suitability of a Tobit model
for ecolbs?

```{r}
# look at a sample of the data
head(apple_data$ecolbs, n=10)
````

The variable ecolbs does not have a continous distribution. This type of distribution violates the latent variable formulation underlying the Tobit model.

## Part (iii)
Estimate a Tobit model for ecolbs with ecoprc, regprc, faminc, and hhsize as explanatory variables. Which variables are significant at the 1% level?

```{r}
# estimate a Tobit model
library(AER)
tobit_model1<-tobit(ecolbs~ecoprc+regprc+faminc+hhsize, data=apple_data)
stargazer(tobit_model1, type="text")

````

The variables significant at the 1% level include: 

## Part (iv)
Are faminc and hhsize jointly significant?

```{r}


````

## Part (v)
Are the signs of the coeffficient on the price variables from part(iii) what you expect? Explain

```{r}


````

## Part (vi)
Test the hypothesis.

```{r}


````

## Part (vii)
Obtain the estimates of E(ecolbs|x) for all observations in the sample.

```{r}


````

## Part (viii)
Compute the squared correlation between ecolbs and estimated ecolbs.

```{r}


````

## Part (ix)
Now, estimate a linear model for ecolbs using the same explanatory variables from part (iii).

```{r}


````

## Part (x)
Evaluate the following statement: "Because the R-squared from the Tobit model is so small, the estimated price effects are probably inconsisent."

```{r}


````

# Problem C17.10
Use the data in SMOKE.RAW for this exercise. 

## Part (i)
How many people in the sample do not smoke at all? what fraction of people claim to smoke 20 cigarettes a day? Why do you think there is a pile up of people at 20 cigarettes?

```{r}


````

## Part (ii)
Does cigs seem a good candidate for havinf a condidtional Poisson distribution?

```{r}


````

## Part (iii)
Estimate a Poisson regression model?

```{r}


````

## Part (iv)
Using the maximum likelihood standard errors, are the price and income variables sististically significant at the 5% level?

```{r}


````

## Part (v)
Obtain the estimate of the variance. What is the estimated standard deviation? How should you adjust the standard errors from part (iv)?

```{r}


````

## Part (vi)
Using the adjusted standard errors, are the price and income elasticities now statistically different from zero? Explain.

```{r}


````

## Part (vii)
Are the education and age variables significant using the more robust standard errors? How do interpret the coefficient on educ?

```{r}


````

## Part (viii)
obtained the fitted values from the poisson regression model. Find the minimum and maximum values.

```{r}


````

## Part (ix)
Using the fitted values from part (viii), obtain the squared correlation coefficient between esimated y and y.

```{r}


````

## Part (x)
Estimate a linea model for cigs by OLS, using the explanatory variables

```{r}


````
