---
title: "problemThree"
author: "Daniel"
date: "April 8, 2015"
output: html_document
---

# Problem C15.3
Use the data in CARD.RAW for this exercise.  

## Subsection (i)
The equation we estimated in Example 15.4 can be written as  

log(wage) = beta0 + beta1educ + beta2exper + ... + u,

where the other explanatory variables are listed in Table 15.1. In order for IV to be consistent, the IV for educ, nearc4, must be uncorrelated with u. Could nearc4 be correlated with things in the error term, such as unobserved ability? Explain.  

It's possible for proximity to college to be correlated with things in the error term such as unobserrved ability. One could argue that the presence of a college can motivate an individual to work harder in school so that they can one day make to college themselves.


## Subsection (ii)
For a subsample of the men in the data set, an IQ score is available. Regress IQ on
nearc4 to check whether average IQ scores vary by whether the man grew up near
a four-year college. What do you conclude?

```{r}
# read the data into R
install.packages("haven")
library(haven)
card_data <- read_dta("CARD.dta")
str(card_data)

# Regress IQ on nearc4
lm_model1 <- lm(iq~nearc4, data = card_data)

# Put the model in a tabular format
install.packages("stargazer")
library(stargazer)
stargazer(lm_model1, type = "text", title = "Regression of IQ on Proximity to College")
```

The model shows that the predicted IQ is 2.6 points higher if the man grew up near a four-year college. This estimate is statistically significant at a 0.01 level of confidence.


## Subsection (iii)
Now, regress IQ on nearc4, smsa66, and the 1966 regional dummy variables
reg662, ... ,reg669. Are IQ and nearc4 related after the geographic dummy variables
have been partialled out? Reconcile this with your findings from part (ii).

```{r}
# Regress IQ on nearc4, smsa66, and reg662, ... ,reg669.
lm_model2 <- lm(iq~nearc4 + smsa66 + reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668 + reg669, data = card_data)

stargazer(lm_model2, type = "text", title = "Regression of IQ on More Variables")
```

The model shows that the relationship between IQ and nearc4 is much weaker and statistically insignificant. This shows that when we control for region and environment while growing up, there is no apparent link between IQ score and living near a four-year college. 


## Subsection (iv)
 From parts (ii) and (iii), what do you conclude about the importance of controlling
for smsa66 and the 1966 regional dummies in the log(wage) equation?  

It is important because it allows control for differences in access to colleges that
might also be correlated with ability. 


# Problem C15.5
Use the data in CARD.RAW for this exercise.

## Subsection (i)
In Table 15.1, the difference between the IV and OLS estimates of the return to education is economically important. Obtain the reduced form residuals, v2, from the reduced form regression educ on nearc4, exper, expersq, black, smsa, south, smsa66, reg662, ..., reg669 ---see Table15.1. Use these to test whether educ is exogenous; that is, determine if the difference between OLS and IV is statistically significant.

```{r}
# The reduced form is obtained as follows
lm_model3 <- lm(educ ~ nearc4 + exper + expersq + black + smsa + south + smsa66 + reg662, + reg663 + reg664 + reg665 + reg666 + reg667 + reg668 + reg669, data = card_data)

# Obtain the residuals
v_residuals <- resid(lm_model3)

# Add the residuals to the original equation
```


## Subsection (ii)
Estimate the equation by 2SLS, adding nearc2 as an instrument. Does the coefficient on educ change much?
```{r}
# Estimate the equation by 2SLS
install.packages("AER")
library(AER)

fit_ivreg <- ivreg()

```


## Subsection (iii)
Test the single overidentifying restriction from part (ii).



# Problem C15.8
Use the data in 401KSUBS.RAW for this exercise. 

## Subsection (i)
Estimate the equation by OLS and discuss the estimated effect of p401k.
```{r}
# Read the data into R
ksubs_data <- read_dta("401KSUBS.dta")

# Fit a linear model by OLS
fit_lm <- lm(pira ~ p401k + inc + incsq + age, data = ksubs_data)

# Tabulate the model results
stargazer(fit_lm, type = "text", title = "Model Results using OLS")
```

The model shows that having a 401(k) plan increases the probability of having an individual retirement account by 0.052.   

## Subsection (ii)
For the purposes of estimating the ceteris paribus tradeoff between participation in two different types of retirement savings plans, what might be a problem with ordinary least squares?

The model in part(i) does not account for the fact that different people may have different saving preferences. People who are savers tend to have both a 401(k) plan as well as an IRA. This makes the error term positively correlated with p401k. 


## Subsection (iii)
The variable e401k is a binary variable equal to one if a worker is eligible to participate in a 401(k) plan. Explain what is required for e401k to be a valid IV for p401k. Do these assumptions seem reasonable?  

Condition 1: e401k has to be correlated with p401k; not surprisingly, this is not an
issue, as being eligible for a 401(k) plan is, by definition, necessary for participation. (The
regression in part (iv) verifies that they are strongly positively correlated.) The more difficult
issue is whether e401k can be taken as exogenous in the structural model. In other words, is
being eligible for a 401(k) correlated with unobserved taste for saving? If we think workers that
like to save for retirement will match up with employers that provide vehicles for retirement
saving, then u and e401k would be positively correlated. Certainly we think that e401k is less
correlated with u than is p401k. But remember, this alone is not enough to ensure that the IV
estimator has less asymptotic bias than the OLS estimator; see page 493. 


## Subsection (iv)
Estimate the reduced form for p401k and verify that e401k has significant partial correlation with p401k. Since the reduced form is also a linear probability model, use a heteroskedasticity-robust standard error.



## Subsection (v)
Now, estimate the structural equation by IV and compare the estimate of b1 with the
OLS estimate. Again, you should obtain heteroskedasticity-robust standard errors.


## Subsection (vi)
Test the null hypothesis that p401k is in fact exogenous, using a heteroskedasticityrobust
test.  


# Problem C15.9
The purpose of this exercise is to compare the estimates and standard errors obtained by
correctly using 2SLS with those obtained using inappropriate procedures. Use the data
file WAGE2.RAW.

## Subsection (i)
Use a 2SLS routine to estimate the equation  

log(wage) 5 b0 1 b1educ 1 b2exper 1 b3tenure 1 b4black 1 u,  

where sibs is the IV for educ. Report the results in the usual form.


## Subsection (ii)
Now, manually carry out 2SLS. That is, first regress educi
 on sibsi
, experi
, tenurei
,
and blacki
 and obtain the fitted values, educi
, i 5 1, ..., n. Then, run the second
stage regression log(wagei
) on educ i
, experi
, tenurei
, and blacki
, i 5 1, ..., n. Verify
that the b ˆ
j are identical to those obtained from part (i), but that the standard errors
are somewhat different. The standard errors obtained from the second stage regression
when manually carrying out 2SLS are generally inappropriate.


## Subsection (iii)
Now, use the following two-step procedure, which generally yields inconsistent
parameter estimates of the bj
, and not just inconsistent standard errors. In step
one, regress educi
 on sibsi
 only and obtain the fitted values, say educi
. (Note that
this is an incorrect first stage regression.) Then, in the second step, run the regression
of log(wagei
) on educi
, experi
, tenurei
, and blacki
, i 5 1, …, n. How does the
estimate from this incorrect, two-step procedure compare with the correct 2SLS
estimate of the return to education?



# Problem C15.10
Use the data in HTV.RAW for this exercise.

## Subsection (i)
Run a simple OLS regression of log(wage) on educ. Without controlling for other
factors, what is the 95% confidence interval for the return to another year of
education?

## Subsection (ii)
The variable ctuit, in thousands of dollars, is the change in college tuition facing
students from age 17 to age 18. Show that educ and ctuit are essentially uncorrelated.
What does this say about ctuit as a possible IV for educ in a simple regression
analysis?

## Subsection (iii)
Now, add to the simple regression model in part (i) a quadratic in experience and a
full set of regional dummy variables for current residence and residence at age 18.
Also include the urban indicators for current and age 18 residences. What is the
estimated return to a year of education?

## Subsection (iv)
Again using ctuit as a potential IV for educ, estimate the reduced form for educ.
[Naturally, the reduced form for educ now includes the explanatory variables in
part (iii).] Show that ctuit is now statistically significant in the reduced form for
educ.  

## Subsectioin (v)
Estimate the model from part (iii) by IV, using ctuit as an IV for educ. How does
the confidence interval for the return to education compare with the OLS CI from
part (iii)?


## Subsection (vi)
Do you think the IV procedure from part (v) is convincing?